Operational Feature Selection in Gaussian Mixture Models

### 2016-09-27 paper.tex

>>>>

\section{Datasets}
\label{sec:datasets}
Experimentation has been conducted on two different datasets. The first dataset, called Aisa, is an airborne hyperspectral dataset and second, called Potsdam, is a remote sensing dataset made of heterogeneous spatial features.

    \subsection{Aisa dataset}
    \label{sec:aisa-dataset}
    The Aisa dataset has been acquired by the AISA Eagle sensor during a flight campaign over Heves, Hungary. It contains 252 bands ranging from 395 to 975 nm. 16 classes have been defined for a total of 361,971 referenced pixels. Figure~\ref{fig:aisa} shows a colored composition of the multispectral image and the shapefile representing groundtruth and Table \ref{tab:aisa} presents the repartition of the various classes.

    \subsection{Potsdam dataset}
    \label{sec:pots-dataset}

    This second dataset is built from a dataset of remote sensing images distributed by the International Society for Photogrammetry and Remote Sensing (ISPRS)\footnote{\url{http://www2.isprs.org/commissions/comm3/wg4/2d-sem-label-potsdam.html}}. The dataset is composed of aerial images of the urban area of Potsdam. The area is cut into 38 patches of 6000x6000 pixels with a resolution of 5cm by pixel and 4 channels are available: Red, Blue, Green and Infrared (RGBIR). A Digital Surface Model with same resolution is also provided and a so-called normalized DSM representing the height above ground. The groundtruth for 24 tiles are provided with 6 classes: Low vegetation, High vegetation, Impervious surfaces, Buildings, Cars, Clutter. Table~\ref{tab:potsdam} summarizes the number of samples of each class in the three images used during experimentations and Figure~\ref{fig:potsdam-expl} the data itself.

    \begin{figure}[!t]
        \centering
        \begin{tabular}{ccc}
            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_5_11_RGB.jpg} &
            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_5_12_RGB.jpg} &
            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_3_10_RGB.jpg} \\
            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_5_11_label.jpg} &
            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_5_12_label.jpg} &
            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_3_10_label.jpg} \\
            \includegraphics[width=0.3\columnwidth]{Fig/dsm_potsdam_05_11_normalized_lastools.jpg} &
            \includegraphics[width=0.3\columnwidth]{Fig/dsm_potsdam_05_12_normalized_lastools.jpg} &
            \includegraphics[width=0.3\columnwidth]{Fig/dsm_potsdam_03_10_normalized_lastools.jpg} \\
            {\bfseries{(a)}} & {\bfseries{(b)}}  & {\bfseries{(c)}}\\
        \end{tabular}
        \caption{From top to bottom, image RGB, reference and normalized DSM of: {\bfseries{(a)}} tile 5\_11, {\bfseries{(b)}} tile 5\_12 and {\bfseries{(c)}} tile 3\_10.\label{fig:potsdam-expl}}
    \end{figure}

    \begin{table}[!t]
        \centering
        \caption{Repartition of classes in Aisa dataset.\label{tab:potsdam}}
        \begin{tabular}[b]{lrrr}\hline
          Class & Number of samples & Number of samples & Number of samples \\
                & in 5\_11          & in 5\_12          & in 3\_10 \\
          \hline
          Clutter             & 1,078,611  & 812,038    & 1,890,467 \\
          Trees               & 4,493,295  & 2,132,368  & 8,780,245 \\
          Cars                & 900,076    & 1,101,541  & 434,615 \\
          Buildings           & 13,469,575 & 17,501,421 & 5,128,149 \\
          Low vegetation      & 4,718,219  & 3,210,596  & 11,428,326 \\
          Impervious surfaces & 11,340,224 & 11,242,036 & 8,338,198 \\
          \hline
        \end{tabular}
    \end{table}

    In the experiment, the following features are computed using the RGBIR images similarly to \cite{tuia2015multiclass}:
    \begin{itemize}
        \item Radiometric indexes: NDVI, TNDVI, RVI, SAVI, TSAVI, MSAVI, MSAVI2, GEMI, IPVI, NDWI2, NDTI, RI, CI, BI, BI2 (15 features) \footnote{\url{https://www.orfeo-toolbox.org//Applications/RadiometricIndices.html}};
        \item Morphological profile with reconstruction of each band with a disk of radius 5, 9, 13, 17, 21, 25, 29, 33, 37 and 41 (80 features);
        \item Attribute profile of each band with area as attribute and 1000, 2000, 5000, 10000 and 15000 as thresholds (40 features);
        \item Attribute profile of each band with diagonal of bounding box as attribute and 100, 200, 500, 1000 and 20000 as thresholds (40 features);
        \item Textural features with neighborhood of 19x19 pixels: mean, standard deviation, range and entropy (16 features).
    \end{itemize}
    The normalized DSM and the raw RGBIR image are added to these 191 features and then then stacked to create a new image with 196 bands. The training is made with tile 5\_11 and test with two tiles: 5\_12 which very similar with tile 5\_11 and 3\_10 which a further area a bit less urban and more industrial.
